---

---

Part 1: Tidy and get ready for Students dataset 
want NA in both waves or NA+Dont Know or Dont Know+NA or Dont Know in both 

```{r}
xx1=is.na(Student_race_all_result$Wave1)&is.na(Student_race_all_result$Wave2)
subset1=Student_race_all_result[xx1,]   
'This is the response from those have NA values in both Waves'
```


```{r}
xx2=is.na(Student_race_all_result$Wave1)
xx2
subset2=Student_race_all_result[xx2,] 
subset2=subset2[subset2$Wave2=='Dont_Know',]
subset2=unique(subset2)
subset2                                  
'This is the subset of students whose response is NA in Wave1 and Dont Know in Wave2'
```

```{r}
xx3=is.na(Student_race_all_result$Wave2)
xx3
subset3=Student_race_all_result[xx3,] 
subset3=subset3[subset3$Wave1=='Dont_Know',]
subset3=unique(subset3)
subset3
'This is the subset if students whose response is Dont Know in Wave1 and NA in Wave2'
```

```{r}

subset4=Student_race_all_result[Student_race_all_result$Wave1=='Dont_Know'&Student_race_all_result$Wave2=='Dont_Know',] 

subset4=unique(subset4)
subset4
```
Then we can combine these four situations together and form the final dataset for students with missing responses.

```{r}
Need_TR_Students=rbind(subset1,subset2,subset3,subset4)
Need_TR_Students
'Then for final check, we want to remove any response with a missing Student ID'

logic=is.na(Need_TR_Students$Student_ID)
sum(logic)
Need_TR_Students=Need_TR_Students[!logic,]
Need_TR_Students
'Then this is the final version for students who need TR responses to help us identify their races'
```




Part 2: Organize TR Responses and get ready for TR dataset

```{r}
sos.wave1_2
'begin with TR, teachers responses on students'

sos.wave1_2_TR<-select(sos.wave1_2,STUDENT_ID,starts_with('TR_Race')) %>%
  select(STUDENT_ID,ends_with('W2')) %>%
  rename_with(~str_remove(., '_W2')) %>%
  mutate(Wave = 2)



sos.wave1_2_TR_long <- select(sos.wave1_2, STUDENT_ID, starts_with('TR_Race')) %>%
  select(STUDENT_ID,ends_with('W1')) %>%
rename_with(~str_remove(., '_W1')) %>%
mutate(Wave = 1) %>%
bind_rows(sos.wave1_2_TR)



sos.wave1_2_TR_long<-unique(sos.wave1_2_TR_long)


sos.wave1_2_TR_long
```



```{r}
colnames(sos.wave1_2_TR_long)<-c('Student_ID','Asian','Black','Haitian','Hispanic','Native','Pac_Island','White','Other','Other_Specify','Wave')

sos.wave1_2_TR_long

'I observe that there are only a few cases which teachers identify students as Haitian but did not label them as Haitian in the survey'
'So need to make one more step of transformation and tidy'

sos.wave1_2_TR_long<-sos.wave1_2_TR_long %>%
  mutate(Haitian=ifelse(Other_Specify=='Haitian',1,0))
sos.wave1_2_TR_long
'Then this is the final version of teachers response on students dataset'
'We will start applying different approach through the catacode function'
```


Part 3: Utilize the Multiple Method

```{r}
TR_race_MU<-catacode(sos.wave1_2_TR_long, id = Student_ID, approach = "multiple", Asian:Other, time=Wave,new.name = "Race_Ethnicity")

TR_race_MU
```

Then we want to match teachers responses with those whose self-identifications are missing

```{r}
Need_TR_Students

Merge_Mu=merge(x=TR_race_MU,y=Need_TR_Students,by='Student_ID')
Merge_Mu
```
300 among 321 students find corresponding TR responses on them.


Part 4: Summarize the result
```{r}
Merge_MU_Summary<-Merge_Mu %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Merge_MU_Summary

```

Part 5: Utilize Priority Approach
```{r}
TR_race_P<-catacode(sos.wave1_2_TR_long, id = Student_ID, approach = "priority", Asian:Other,priority = c('Hispanic','Haitian'), time=Wave,new.name = "Race_Ethnicity")

TR_race_P
```

Merge the responses
```{r}
Merge_P=merge(x=TR_race_P,y=Need_TR_Students,by='Student_ID')
Merge_P
```

Part 6: Summarize the result
```{r}
Merge_P_Summary<-Merge_P %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Merge_P_Summary

```


Part 7: Utilize the Mode Approach

```{r}
TR_race_M<-catacode(sos.wave1_2_TR_long, id = Student_ID, approach = "mode", Asian:Other,priority = c('Hispanic','Haitian'), time=Wave,new.name = "Race_Ethnicity")

colnames(TR_race_M)=c('Student_ID','Race_Ethnicity')
TR_race_M
```
Then merge
```{r}
Merge_M=merge(x=TR_race_M,y=Need_TR_Students,by='Student_ID')
Merge_M
```

Part 8: Summarize the result
```{r}
Merge_M_Summary<-Merge_M %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Merge_M_Summary
```


Comparison
```{r}

Merge_MU_Summary

Merge_P_Summary


Merge_M_Summary
```

