'1) Import Data into R


```{r}
library(haven)
library(tidyverse)
```

```{r}
sos.wave1_2 <- read_sav("SOS_Wave1_2.sav")
sos.wave1_2
```
Start with analyzing teacher's response
```{r}
sos.wave1_2_T<-select(sos.wave1_2,TEACHER_ID,starts_with('T_Race')) %>%
  select(TEACHER_ID,ends_with('W2')) %>%
  rename_with(~str_remove(., '_W2')) %>%
  mutate(Wave = 2)



sos.wave1_2_T_long <- select(sos.wave1_2, TEACHER_ID, starts_with('T_Race')) %>%
  select(TEACHER_ID,ends_with('W1')) %>%
rename_with(~str_remove(., '_W1')) %>%
mutate(Wave = 1) %>%
bind_rows(sos.wave1_2_T)



sos.wave1_2_T_long<-unique(sos.wave1_2_T_long)
"I observe that lots of participants sumbit multiple responses"

sos.wave1_2_T_long
"This is the final version for our long dataset"
```

'Then check T_Race_Other-Specificy this column, it seems like all teacher respodents correctly identify their races even though they put speicfic race other answers in the survey'

```{r}



colnames(sos.wave1_2_T_long)<-c('Teacher_ID','Asian','Black','Native','Pac_Island','Haitian','Hispanic','White','Other','Other_Specify','Wave')

sos.wave1_2_T_long
```
Part 1----Teacher's Response Multiple Approach

```{r}
Teacher_race_MU<-catacode(sos.wave1_2_T_long, id = Teacher_ID, approach = "multiple", Asian:Other, time=Wave,new.name = "Race_Ethnicity")
'This shows the Race_Ethnicity for each participant based on multiple measurement'

Teacher_race_MU

```

'Generate visualizations based on the result'
```{r}
Teacher_race_MU_Summary<-Teacher_race_MU %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Teacher_race_MU_Summary

Teacher_race_MU_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Teacher_race_MU_Summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,50))

Teacher_race_MU_plot
```

Part 2----Teacher's Response Priority Approach
```{r}
Teacher_race_P<-catacode(sos.wave1_2_T_long, id = Teacher_ID, approach = "priority", Asian:Other,priority = c('Hispanic','Haitian'), time=Wave,new.name = "Race_Ethnicity")


Teacher_race_P
```
'Generate visualizations based on the result'
```{r}
Teacher_race_P_Summary<-Teacher_race_P %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Teacher_race_P_Summary

Teacher_race_P_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Teacher_race_P_Summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,50))

Teacher_race_P_plot
```

Part 3---Teacher's Response Mode Approach
```{r}
Teacher_race_M<-catacode(sos.wave1_2_T_long, id = Teacher_ID, approach = "mode", Asian:Other,priority = c('Hispanic','Haitian'), time=Wave,new.name = "Race_Ethnicity")


colnames(Teacher_race_M)<-c('Teacher_ID','Race_Ethnicity')
```

'Generate visualizations based on the result'
```{r}
Teacher_race_M_Summary<-Teacher_race_M %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Teacher_race_M_Summary

Teacher_race_M_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Teacher_race_M_Summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,50))

Teacher_race_M_plot
```
Addtional Comparison: Compare the result, using three different approaches
```{r}
Teacher_race_MU_Summary2<-Teacher_race_MU_Summary%>%
  mutate(Approach='Multiple')

Teacher_race_P_Summary2<-Teacher_race_P_Summary%>%
  mutate(Approach='Priority')

Teacher_race_M_Summary2<-Teacher_race_M_Summary%>%
  mutate(Approach='Mode')

Teacher_combined=rbind(Teacher_race_MU_Summary2,Teacher_race_P_Summary2,Teacher_race_M_Summary2)
Teacher_combined
```

```{r}
Teacher_combined_graph=ggplot(Teacher_combined,aes(x=Race_Ethnicity,y=count))+geom_bar(stat = 'identity',aes(fill=Race_Ethnicity))+facet_wrap(.~Approach)+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,50))
Teacher_combined_graph


```




Part 4---Students Response Multiple Approach

Start with subtracting students' responses
```{r}
sos.wave1_2_ST<-select(sos.wave1_2,STUDENT_ID,starts_with('ST_Race')) %>%
  select(STUDENT_ID,ends_with('W2')) %>%
  rename_with(~str_remove(., '_W2')) %>%
  mutate(Wave = 2)



sos.wave1_2_ST_long <- select(sos.wave1_2, STUDENT_ID, starts_with('ST_Race')) %>%
  select(STUDENT_ID,ends_with('W1')) %>%
rename_with(~str_remove(., '_W1')) %>%
mutate(Wave = 1) %>%
bind_rows(sos.wave1_2_ST)



sos.wave1_2_ST_long<-unique(sos.wave1_2_ST_long)
"I observe that lots of participants sumbit multiple responses"

colnames(sos.wave1_2_ST_long)<-c('Student_ID','Asian','Black','Native','Other_Specify','Other','Pac_Island','White','Dont_Know','Wave')

sos.wave1_2_ST_long

'Then,for double check, I checked students responses on Race_Other_Specify this column. Considering their young age, they may encounter some difficulties and i want to make sure they didnt put any provided race values in this column'


'For Hispanic ones, they are labelled as 0 in Others Specify, for Haitian ones, they are labelled as 1, for Whitrs, they are labelled as 2'
 sos.wave1_2_ST_long<-sos.wave1_2_ST_long %>%
  mutate(Other_Specify=ifelse(Other_Specify=='hispanic'|Other_Specify=='Spanish'|Other_Specify=='spanish',0,Other_Specify)) %>%
  mutate(Other_Specify=ifelse(Other_Specify=='haitian'|Other_Specify=='Haitian'|Other_Specify=='haiitan',1,Other_Specify)) %>%
   mutate(Other_Specify=ifelse(Other_Specify=='white'|Other_Specify=='White',2,Other_Specify))
 



'Then make sure these values are reflected in their original columns'
'because in the original survey, for students verion, Hispanic and Haitian is not provided, so I create two new columns to record them'
sos.wave1_2_ST_long2<-sos.wave1_2_ST_long %>%
  mutate(Hispanic=ifelse(Other_Specify==0,1,0)) %>%
  mutate(Haitian=ifelse(Other_Specify==1,1,0)) %>%
  mutate(White=ifelse(Other_Specify==2,1,0)) %>%
  select(-Other_Specify)

sos.wave1_2_ST_long2
'This is the final version for students responses dataset'
```


Students Response Multiple Approach
```{r}
Student_race_MU<-catacode(sos.wave1_2_ST_long2, id = Student_ID, approach = "multiple", c(Asian:Dont_Know,Hispanic:Haitian), new.name = "Race_Ethnicity")

Student_race_MU


Student_race_MU_summary<-Student_race_MU %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Student_race_MU_summary

Student_race_MU_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Student_race_MU_summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,80))

Student_race_MU_plot
```

Part 5: 
Students Response Priority Approach
```{r}
Student_race_P<-catacode(sos.wave1_2_ST_long2, id = Student_ID, approach = "priority", c(Asian:Dont_Know,Hispanic:Haitian),priority = c('Hispanic','Haitian'), new.name = "Race_Ethnicity")

Student_race_P


Student_race_P_summary<-Student_race_P %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Student_race_P_summary

Student_race_P_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Student_race_P_summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,80))

Student_race_P_plot
```

Part 6: Students rsponse Priority Approach
```{r}
Student_race_M<-catacode(sos.wave1_2_ST_long2, id = Student_ID, approach = "mode", c(Asian:Dont_Know,Hispanic:Haitian),priority = c('Hispanic','Haitian'), new.name = "Race_Ethnicity")

colnames(Student_race_M)<-c('Student_ID','Race_Ethnicity')


Student_race_M_summary<-Student_race_M %>%
  group_by(Race_Ethnicity) %>%
  summarize(count=n())

Student_race_M_summary

Student_race_M_plot<-ggplot(aes(x=Race_Ethnicity,y=count),data=Student_race_M_summary)+geom_bar(stat='identity',aes(fill=Race_Ethnicity))+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,80))

Student_race_M_plot
```

Additional Comparision of students responses
```{r}
Student_race_MU_summary
Student_race_P_summary
Student_race_M_summary

Student_race_MU_Summary2<-Student_race_MU_summary%>%
  mutate(Approach='Multiple')

Student_race_P_Summary2<-Student_race_P_summary%>%
  mutate(Approach='Priority')

Student_race_M_Summary2<-Student_race_M_summary%>%
  mutate(Approach='Mode')

Student_combined=rbind(Student_race_MU_Summary2,Student_race_P_Summary2,Student_race_M_Summary2)
Student_combined
```

```{r}
Student_combined_graph=ggplot(Student_combined,aes(x=Race_Ethnicity,y=count))+geom_bar(stat = 'identity',aes(fill=Race_Ethnicity))+facet_wrap(.~Approach)+geom_text(aes(label=count,vjust=-4))+scale_y_continuous(limits = c(0,50))
Student_combined_graph

```



Part 7: Distribution of Teachers whose answers are different across waves
```{r}
Teacher_race_all<-catacode(sos.wave1_2_T_long, id = Teacher_ID, approach = "all", Asian:Other, time=Wave,new.name = "Race_Ethnicity")
Teacher_race_all_result<-pivot_wider(data=Teacher_race_all,names_from = Wave,values_from = Race_Ethnicity)
'Change the summary table from long to wide format'
colnames(Teacher_race_all_result)=c('Teacher_ID','Wave1','Wave2')
Teacher_race_all_result<-Teacher_race_all_result%>%
  mutate(same=ifelse(Wave1==Wave2,1,0))
Teacher_race_all_result
sum(na.omit(Teacher_race_all_result$same))
```
25 among 68 teachers responses are the same across waves



Part 8: Distribution of Students whose answers are different across waves
```{r}
Student_race_all<-catacode(sos.wave1_2_ST_long2, id = Student_ID, approach = "all", c(Asian:Dont_Know,Hispanic:Haitian),time=Wave, new.name = "Race_Ethnicity")
Student_race_all

Student_race_all_result<-pivot_wider(data=Student_race_all,names_from = Wave,values_from = Race_Ethnicity)
'Change the summary table from long to wide format'
colnames(Student_race_all_result)=c('Student_ID','Wave1','Wave2')

Student_race_all_result

Student_race_all_result<-Student_race_all_result%>%
  mutate(same=ifelse(Wave1==Wave2,1,0))
Student_race_all_result
sum(na.omit(Student_race_all_result$same))
```
52 among 504 students have the exactly same responses across waves
